Privacy-Respecting Type Error Telemetry at Scale

### 2023-10-12 programming23/paper.tex
fig 6 caption
delete comment
power law check
except for lines, they all have p>0.2 which is what Clauset et al said was good

lines has p=0.1

did not add prose to the paper, this doesn't seem important after all
but maybe we'll have space to include in the final version

\begin{figure}[t]\centering
  %% TODO bigger text
  %% TODO weird artifact (|) in x-min label
  %% code/row-distribution.rkt
  \includegraphics[width=\columnwidth]{img/row-distribution.pdf}
  \begin{tabular}{ll}
    Reason for sending:
    &
    \begin{tabular}[t]{r@{~~}l@{~~}r}
      996,164 & due to keystrokes  & [\pct{66.20}] \\
      508,572 & due to module switch & [\pct{33.80}]
    \end{tabular}
  \end{tabular}
  \caption{Telemetry records per hour. Each tick on the $x$-axis marks the start of a new day in California. Shaded ranges correspond to weekends.}
  \label{f:records-per-hour}
\end{figure}

@@

Three important characteristics of the recorded data are the size of the codebases,
the length of the sessions, and the number of type
errors.
We discuss these in turn.
% Commenting these out because that's literally the rest of the
% subsection. Say anything you want from here over there!
% Codebase size and session length have long-tailed
% distributions\todo{DID YOU DO A STATISTICAL TEST.}
% with many small items and few very large ones.
% %% TODO check %% follow power-law distributions~\cite{csn-siamr-2009} 
% %% https://aaronclauset.github.io/powerlaws/
% %% bin version, is this relevant? sounds like no from abtract https://arxiv.org/pdf/1208.3524.pdf
% The number of type analysis errors is somewhat low~(explained in the next
% section), but sufficient to proceed with futher analysis.

@@

There is a huge amount of variation across projects.
The largest ones have over 50,000 lines of code
while the smallest have merely 1 line of code.
Unsurprisingly, these wide-ranging numbers come with large standard
deviations.
The median values are more reasonable, with roughly 3,000 lines of code
and 1,000 lines in edit ranges.

@@

\begin{table}[t]\centering
  \caption{Session size in seconds and in number of records.}
  \label{t:session-size}
  %% NOTE max time = 1 million seconds = 15 days ...  +100 are over 2 days
  \begin{tabular}{lrrrrl}
    ~               & Mean & Stddev & Median & \ptile{99} & Distribution \\
    Time Span (sec) & 3,184 & \stddev{16} & 845 & 35,450 & \includegraphics[width=0.14\columnwidth,valign=M]{img/timespan-distribution.pdf} \\
    Record Count    & 286 & 583         & 138 & 3,302  & \includegraphics[width=0.14\columnwidth,valign=M]{img/event-count-distribution.pdf}
    %% TODO x-axis starts at 1
  \end{tabular}
\end{table}

\Cref{t:session-size} similarly outlines the size of sessions using two metrics:
real time in seconds between the first and last record, and the total number of
records.
The plots focus
on the lower-left fragment of the data, which again has a very long tail.
%% comment on long-tail record counts?
The longest sessions last several days and/or contain thousands of records,
while the shortest last a few milliseconds or include just one record.
% Likewise, both plots in the figure follow a power-law distribution
% (especially for record count).
A median session runs on the order of minutes and consists of a few dozen events.
%% The median time is 14 minutes and the median 
%% The median number of records is 138.
% These numbers are reasonable.

@@

% BEN: double-check, I think this says all you were trying to say.
% "typically" because we COULD have gotten lucky.
% ben: I'd written "immediatly introduces" to rule out getting lucky. Deleted "immediately" and went with your version.
The counts are imprecise because of telemetry granularity.
If a creator removes one error but then
introduces another on a different line, this typically manifests as
a \keepsym{} rather than one \addsym{} and one \dropsym{}.

@@

Fortunately, the data rarely contains too-complex errors.
None of these errors appear in edit ranges~(\cref{t:type-error-count}).
For the specific case of \code{CodeTooComplex}, our telemetry tracked project-wide
counts and found only 26 errors, which were spread across eleven
records in three sessions.
Note that all three sessions \emph{could} have been from the same codebase.
% BEN: I see no reason for this speculation. The only important thing
% is the fact that this could have been all the same codebase.
%It is entirely possible that they all came from one session that
%deliberately pushed the limits of type analysis.
%% Whatever the reason, the overall number is low enough to suggest that the
%% internal limits are not posing barriers to typechecker adoption.

@@

\begin{figure}[t]\centering
  %% FS plots look similar
  %% - much higher densities, same basic story
  %% - nocheck indeed the highest > NS > S
  %% TODO draw units, bigger font

@@

  \mstrict{}
  \includegraphics[width=\columnwidth]{img/error-count-strict-row--te-density-diff.pdf}
  \caption{Changes in type error density (errors / lines) over time (seconds).}
  \label{f:error-density}
\end{figure}

>>>>

\begin{figure}[t]\centering
  %% TODO weird artifact (|) in x-min label
  %% code/row-distribution.rkt
  \includegraphics[width=\columnwidth]{img/row-distribution.pdf}
  \begin{tabular}{ll}
    Reason for sending:
    &
    \begin{tabular}[t]{r@{~~}l@{~~}r}
      996,164 & due to keystrokes  & [\pct{66.20}] \\
      508,572 & due to module switch & [\pct{33.80}]
    \end{tabular}
  \end{tabular}
  \caption{Telemetry records per hour. Each tick on the $x$-axis marks the start of a new day in California. Shaded ranges correspond to weekends.}
  \label{f:records-per-hour}
\end{figure}

@@

Three important characteristics of the recorded data are the size of the codebases,
the length of the sessions, and the number of type
errors.
We discuss these in turn.

@@

There is a huge amount of variation across projects.
The largest ones have over 50,000 lines of code
while the smallest have zero lines of code.
%% TODO filter the zero lines? what happens in that data?? run a query
Unsurprisingly, these wide-ranging numbers come with large standard
deviations.
The median values are more reasonable, with roughly 3,000 lines of code
and 1,000 lines in edit ranges.

%% power law test~\cite{csn-siamr-2009,g-jss-2015}
%%  lines     xmin = 2 alpha = 1.46 p = 0.1 (low!)
%%  editrange xmin = 1 alpha = 1.48 p = 0.73

@@

\begin{table}[t]\centering
  \caption{Session size in seconds and in number of records.}
  \label{t:session-size}
  %% NOTE max time = 1 million seconds = 15 days ...  +100 are over 2 days
  \begin{tabular}{lrrrrl}
    ~               & Mean & Stddev & Median & \ptile{99} & Distribution \\
    Time Span (sec) & 3,184 & \stddev{16} & 845 & 35,450 & \includegraphics[width=0.14\columnwidth,valign=M]{img/timespan-distribution.pdf} \\
    Record Count    & 286 & 583         & 138 & 3,302  & \includegraphics[width=0.14\columnwidth,valign=M]{img/event-count-distribution.pdf}
  \end{tabular}
\end{table}

\Cref{t:session-size} similarly outlines the size of sessions using two metrics:
real time in seconds between the first and last record, and the total number of
records.
The plots focus
on the lower-left fragment of the data, which again has a very long tail.
%% comment on long-tail record counts?
The longest sessions last several days and/or contain thousands of records,
while the shortest last a few milliseconds or include just one record.
%% 0 ms = 1 event
A median session runs on the order of minutes and consists of a few dozen events.

%% power law test~\cite{csn-siamr-2009,g-jss-2015}
%%  timespan    xmin = 1 alpha = 2.88 p = 0.23
%%  event-count xmin = 1 alpha = 1.33 p = 0.86

@@

The counts are imprecise because of telemetry granularity.
If a creator removes one error but then
introduces another on a different line, this typically manifests as
a \keepsym{} rather than one \addsym{} and one \dropsym{}.

@@

Fortunately, the data rarely contains too-complex errors.
None of these errors appear in edit ranges~(\cref{t:type-error-count}).
For the specific case of \code{CodeTooComplex}, our telemetry tracked project-wide
counts and found only 26 errors, which were spread across eleven
records in three sessions.
Note that all three sessions \emph{could} have been from the same codebase.

@@

\begin{figure}[t]\centering
  %% FS plots look similar
  %% - much higher densities, same basic story
  %% - nocheck indeed the highest > NS > S

@@

  \mstrict{}
  \includegraphics[width=\columnwidth]{img/error-count-strict-row--te-density-diff.pdf}
  \caption{Changes in type error density ($\frac{\mbox{\#errors}}{\mbox{\#lines}}$) over time (seconds).}
  \label{f:error-density}
\end{figure}
